---
format:
  gfm:
    default-image-extension: ""
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# econiodatajp

<!-- badges: start -->
<!-- badges: end -->

econiodatajp provides ready-to-use Japanese input-output (IO) tables,
built from data published by the Ministry of Internal Affairs and
Communications ([e-Stat](https://www.e-stat.go.jp/)) and other sources.
Each table is returned as an [econio](https://github.com/UchidaMizuki/econio)
input-output table object, so it works directly with econio's analysis
functions (`io_reclass()`, `io_leontief_inverse()`, and so on).

Tables are archived and rebuilt on demand via
[tarchives](https://github.com/UchidaMizuki/tarchives)/
[targets](https://books.ropensci.org/targets/), so `io_table_get()` (or
`io_table_target()`, for use inside your own `_targets.R` pipeline)
downloads and processes each source file only once per version.

## Installation

You can install the development version of econiodatajp from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("UchidaMizuki/econiodatajp")
```

## Usage

`io_table_get()` covers three shapes of table, selected via `region_type`
and `region_class`. `sector_class` has no default and must always be
specified; `language` defaults to `"ja"` (the original Japanese sector
names) and only accepts `"en"` for the national table.

The national IO table (`region_type = "regional"` and `region_class =
"nation"` are both the default):

```{r}
library(econiodatajp)

io_table_get(year = 2020, sector_class = "medium")
```

Pass `language = "en"` for English sector names:

```{r}
io_table_get(year = 2020, sector_class = "medium", language = "en")
```

One prefecture (`region_class = "pref"` with `region_type = "regional"`
requires `region`, which accepts a numeric code or a `"NN_name"`
fragment); `sector_class` is that prefecture's own sector count, not a
tier shared across prefectures -- use `io_table_available()` to look it
up:

```{r}
io_table_get(region_class = "pref", region = 1, year = 2015, sector_class = "105")
```

Every prefecture at once, as a single table with a region dimension
(`region` isn't used for a `"multiregional"` table -- it always covers
every region in the breakdown at once):

```{r}
io_table_get(
  region_type = "multiregional",
  region_class = "pref",
  year = 2011,
  sector_class = "31"
)
```

Or the official 9-region block breakdown instead (`region_class =
"block"`; published for 1970, 1975, 1980, 1985, 1990, 1995, and 2005 --
use `io_table_available()` to list what's on hand):

```{r}
io_table_get(
  region_type = "multiregional",
  region_class = "block",
  year = 2005,
  sector_class = "12"
)
```

`sector_class` controls the sector classification granularity
(`"basic"`, `"small"`, `"medium"`, `"large"`, or `"template"` for
`region_class = "nation"`, each a direct translation of a real official
Japanese classification tier name). No official source names the sector
granularities of a `region_class = "pref"` or `"block"` table, only
counts them, so `sector_class` there is that count as a string instead:
each prefecture's own count for its `region_type = "regional"` table
(e.g. `"105"` for Hokkaido); `"26"`/`"31"` (2005/2011) for the
`region_class = "pref"` multiregional table; and `"12"`/`"29"`/`"53"` in
2005, or a single year-specific count otherwise (e.g. `"43"` in 1970),
for the `region_class = "block"` multiregional table -- use
`io_table_available()` to look any of these up.
`competitive_import` and `language` are only meaningful for
`region_class = "nation"`.

To declare one of these tables as a target in your own `{targets}`
pipeline instead of fetching it eagerly, use `io_table_target()` with the
same arguments:

``` r
# _targets.R
library(targets)
library(econiodatajp)

list(
  io_table_target(iotable_nation_2020, year = 2020, sector_class = "medium")
)
```
