---
format:
  gfm:
    default-image-extension: ""
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# econiodatajp

<!-- badges: start -->
<!-- badges: end -->

econiodatajp provides ready-to-use Japanese input-output (IO) tables,
built from data published by the Ministry of Internal Affairs and
Communications ([e-Stat](https://www.e-stat.go.jp/)) and other sources.
Each table is returned as an [econio](https://github.com/UchidaMizuki/econio)
input-output table object, so it works directly with econio's analysis
functions (`io_reclass()`, `io_leontief_inverse()`, and so on).

Tables are archived and rebuilt on demand via
[tarchives](https://github.com/UchidaMizuki/tarchives)/
[targets](https://books.ropensci.org/targets/), so `io_table_get()` (or
`io_table_target()`, for use inside your own `_targets.R` pipeline)
downloads and processes each source file only once per version.

## Installation

You can install the development version of econiodatajp from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("UchidaMizuki/econiodatajp")
```

## Usage

`io_table_get()` covers three shapes of table, selected via `region_type`
and `area`.

The national IO table (`region_type = "regional"` is the default; `area`
defaults to `"nation"`). `language` defaults to `"en"` with a one-time
note; pass `language = "ja"` for the original Japanese sector names:

```{r}
library(econiodatajp)

io_table_get(year = 2020, sector_class = "medium")
```

One prefecture (`area` accepts a numeric code or a `"NN_name"` fragment):

```{r}
io_table_get(year = 2015, area = 1)
```

Every prefecture at once, as a single table with a region dimension:

```{r}
io_table_get(year = 2011, region_type = "multiregional")
```

`sector_class` controls the sector classification granularity
(`"basic"`, `"small"`, `"medium"`, `"large"`, or `"template"` for the
national table; a single fixed granularity for prefectural/multiregional
tables). `competitive_import` and `language` are only meaningful for the
national table.

To declare one of these tables as a target in your own `{targets}`
pipeline instead of fetching it eagerly, use `io_table_target()` with the
same arguments:

``` r
# _targets.R
library(targets)
library(econiodatajp)

list(
  io_table_target(iotable_nation_2020, year = 2020, sector_class = "medium")
)
```
